---
title: "EDS 221 Day 3 Comp Session 1"
format: html
editor: visual
execute: 
  warning: false
---

## Conditional Statements

```{r}

```

```{r}
temperature <- 98.2

# if condition isn't met, it returns nothing

if (temperature > 97) {
  print("Temperature warning")
} else 
  print("No temperature warning")
```

Create an if else statement for blue whale lengths. Create a variable blue whale, which has a value of 40 (meters). Add an if else statement that returns "GIANT" if the length exceeds 35, "MEDIUM" if the length is between 20 and 35, or "SMALL" if the length is \<= 20.

```{r}

blue_whale <- 40

if (blue_whale > 35) {
  print("GIANT")
} else if (blue_whale > 20 & blue_whale <= 35) {
  print("MEDIUM")
} else if (blue_whale <= 20) {
  print("SMALL")
}
  

```

## Switch statement

```{r}
housing_type <- "whole"

switch(housing_type, 
       "single" = print("400"),
       "double" = print("750"),
       "whole" = print("1100")
       )
# we can also write this as an if else statement
```

```{r}
if (housing_type == "whole") {
  print("1100")
} else if (housing_type == "double") {
  print("750")
} else if (housing_type == "single") {
  print("400")
}
```

## For Loops
